Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-20 mu/article/article-review7-Anatoli.tex
Sending changes to BMS

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
%\documentclass[preprint,prb,showpacs]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage[pdffitwindow,colorlinks,citecolor={red},linkcolor={blue}]{hyperref}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, one labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first one is by fixing the spin
along a chosen direction, and the resulting SVI is calculated.
Secondly, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt. These speed values are larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimension (2D)
electronic systems; however, most electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we present two 50\% hydrogenated graphene
noncentrosymmetric structures, both presenting a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

Both the Up and the Alt  structures
are noncentrosymmetric, and therefore,
they are good candidates in which SVI can be
induced. In this article, 
we address theoretically the SVI 
by one-photon absorption of linearly polarized light, analyzing
in our structures,
two possible scenarios of practical interest. 
The first case is by fixing the spin of the electrons along $z$, i.e. 
perpendicular 
to the surface plane, with the resulting velocity
directed along the surface of the
structures in the $xy$ plane. 
In the second case we fix the SVI velocity along the
$x$ or $y$ direction, and then, the resulting spin is directed outward of the
$xy$ plane.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{
Top (top panel) and side (bottom panel) views of the Up
      structure along with the 
      Cartesian $x$, $y$, and $z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
{\color{red}Anatoli: we are confused about the primitive hexagonal
  unit cell shown. For instance, according to us, if we take C$_2$ and
apply the red arrow, it will end up in the site of C$_4$, but these C
atoms according to Table \ref{tab:up-unitcell} are not equivalent, or are they?}} 
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{
Top (top panel) and side (bottom panel) views of the Alt
      structure along with the 
      Cartesian $x$, $y$, and$ z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
}
    \label{fig:alt-struc}
\end{figure}

This paper is organized as follows. In Section \ref{sec:theory} we present the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our findings in Section \ref{sec:conclusions}.

@@

The operator that describes the electronic SVI is written as

@@

where $\hat{\mathbf v}=[\hat{\mathbf r},\hat H_0]/i\hbar$ 

@@

@@

k})$ being the energy of the electronic band $n$ at point ${\mathbf k}$ in the

@@

where repeated Cartesians upperscripts  are summed, and 

@@

is the pseudotensor that describes the rate of change of the  PSC process in

@@

K^{\mathrm{ab}}_{nm}(-\mathbf{k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$ which

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system of
coordinates by the defined  polar angle
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and an azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H$_1$ & -0.61516 & -1.77416 &  0.73196 \\
H$_2$ &  0.61518 &  0.35514 &  0.73175 \\
C$_1$ & -0.61516 & -1.77264 & -0.49138 \\
C$_2$ & -0.61516 & -0.35600 & -0.72316 \\
C$_3$ &  0.61516 &  0.35763 & -0.49087 \\
C$_4$ &  0.61518 &  1.77416 & -0.73196 \\
\hline
\end{tabular}
\caption{
Atomic positions in the unit cell of the Up structure shown
in Fig. \ref{fig:up-struc}. 
}
\label{tab:up-unitcell}
\end{table}
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H$_1$ &  -0.61516 &  -1.42140 & \ 1.47237 \\
C$_1$ &  -0.61516 &  -1.73300 & \ 0.39631 \\
C$_2$ & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
C$_3$ & \ 0.61516 & \ 0.42201 &  -0.15814 \\
C$_4$ &  -0.61516 &  -0.37396 &  -0.39632 \\
H$_2$ &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}
\caption{
Atomic positions in the unit cell of the Alt structure
 shown in Fig. \ref{fig:alt-struc}. 
}
\label{tab:alt-unitcell}
\end{table}

@@

We calculated the self-consistent ground state and the Kohn-Sham states within
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used Hartwigsen-
Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented in Eq.
\eqref{eq:mu}.
The convergence parameters for the calculations,  corresponding to the Up and
Alt structures are cutoff energies up to 65\,Ha, resulting in LDA
energy band gaps of 0.084\,eV and 0.718\,eV, respectively, 
and 14452 ${\mathbf k}$ points in the IBZ
where the energy
eigenvalues and matrix elements were calculated;
to
integrate $\mu^{\mathrm{abcd}}(\omega)$ and
$\xi^{\mathrm{a}\mathrm{b}}(\omega)$ the linearized analytic tetrahedron method
(LATM) has been used.\cite{nastosPRB07} 
We neglect the anomalous velocity term
$\hbar(\boldsymbol{\sigma}\times\nabla V)/4m^2c^2$, where $V$ is the crystal
potential, in $\hat{\mathbf v}$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore, $[\hat{\mathbf
v},\hat{\mathbf S}]=0$, and Eq.~\eqref{z.1} reduces to $\hat
K^{\mathrm{a}\mathrm{b}}=\hat v^\mathrm{a} \hat S^\mathrm{b}=\hat S^\mathrm{b}
\hat v^\mathrm{a}$. Finally, the prime in the sum of Eq.~\eqref{eq:mu} is
restricted to quasi-degenerated conduction bands $c$ and $c'$ that are closer
than 30 meV, which  is both typical laser- pulse energy width and the thermal
room-temperature energy level broadening.\cite{nastosPRB07}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  {\it vs.} $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, becoming different from zero at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ becomes different 
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$, parallel to the surface, with their spin
polarized along $z$, perpendicular to the surface. The corresponding speeds
are the following: Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

>>>>

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{soul}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional (2D)
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first,  the spin is fixed
along a chosen direction, and the resulting SVI is calculated.
In the second, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal for
the Alt systems. These speed values are of order of magnitude larger than those of bulk semiconductors
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Graphene, an allotrope of carbon with hexagonal 2D lattice structure,
demonstrates properties such as fractional quantum Hall effect at room
temperature, excellent thermal transport properties, excellent
conductivity\cite{heerscheNat07} and strength \cite{geimNM07, reinaNL08,
novoselov2S07, balandinNL08}, being a perfect platform  for two-dimensional (2D)
electronic systems; however, numerous important  electronic applications are disabled by the
absence of a semiconducting gap. Recent studies demonstrate that a narrow band
gap  can be opened in graphene by applying an electric field,\cite{zhangN09}
reducing the surface area,\cite{hanPRL07} or applying uniaxial
strain.\cite{niACSN08} Another possibility to open the gap is by doping; this
has been successfully achieved using nitrogen,\cite{weiNL2009} 
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} 
noble-metals,\cite{varykhalovPRB10} and hydrogen.\cite{eliasS09, guisingerNL09,
samarakoonACSN10} Depending on the percentage of hydrogenation and spatial
arrangements of the hydrogen-carbon bonds, hydrogenated graphene demonstrates
different structural configurations and a tunable electron gap, as it has been
proven in Ref. \onlinecite{shkrebtiiPSSC12}. 
In this paper, we offer two 50\% hydrogenated graphene
noncentrosymmetric structures, both demonstrating a discernible band gap. The
first one, labelled as the Up structure, also known as
graphone,\cite{gmitraPRL13}  has
hydrogen atoms bonded to the carbon layer only on the upper side of the
structure; we consider here the magnetic isomer of graphone, with the so-called
``chair'' structure  shown in Fig. \ref{fig:up-struc}.
In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

Both the Up and the Alt  structures
are noncentrosymmetric, and therefore,
they are good candidates in which SVI can be
induced. In this article, 
we address theoretically the spin velocity injection
by one-photon absorption of linearly polarized light, analyzing
in our structures
two possible scenarios of practical interest. 
The first case is by fixing the spin of the electrons along $z$, i.e., 
perpendicular 
to the surface plane, with the resulting velocity
directed along the surface of the
structures in the $xy$ plane. 
In the second case we fix the SVI velocity along the
$x$ or $y$ direction, and then, the resulting spin is directed outward of the
$xy$ plane.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{
Top (top panel) and side (bottom panel) views of the Up
      structure along with the 
      Cartesian $x$, $y$, and $z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive hexagonal unit cell is also shown.
{\color{red}Anatoli: we are confused about the primitive hexagonal
  unit cell shown. For instance, according to us, if we take C$_2$ and
apply the red arrow, it will end up in the site of C$_4$, but these C
atoms according to Table \ref{tab:up-unitcell} are not equivalent, or are they? I am absolutely sure that the two atoms are equivalent, this structure, e.g., has the same symmetry as graphane. They might be slightly not equivalent since the calculations have been done for twice-bigger unit cell. And after your question I have left only three numbers after comma, five is too much and they are below the error in the coordinate determination
 }} 
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{
Top (top panel) and side (bottom panel) views of the Alt
      structure along with the 
      Cartesian $x$, $y$, and $z$ directions. The dark (light) spheres
      are the C (H) atoms. 
The primitive rectangular unit cell is also shown.
}
    \label{fig:alt-struc}
\end{figure}

This paper is organized as follows. In Section \ref{sec:theory} we outline the
formalism and the main expressions  that describe PSC and SVI. In Section
\ref{sec:results} we describe the numerical details  and  discuss the
corresponding SVI spectra for the Up and Alt structures. Finally, we summarize
our findings in Section \ref{sec:conclusions}.

@@

To calculate the velocity of the spin injection 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ along the direction $\hat{\mathbf{a}}$, at which the spin moves in a
polarized state along direction $\hat{\mathbf{b}}$,  
we start with the operator that describes the electronic SVI, written as

@@

Here $\hat{\mathbf v}=[\hat{\mathbf r},\hat H_0]/i\hbar$ 

@@

% just for comments to think about, not to introduce the suggestions right away: 
% above we have rho_mn, below rho_cc’
% time variable has disappeared below. And rho_cc’ contains a and b labels,
% which are not Cartesian

@@

k})$ being the energy of the electronic band $n$ and $m$ at point ${\mathbf k}$ in the

@@

where the repeated Cartesians upperscripts  are summed, and 

@@

is the pseudotensor that describes the rate of change of the  PSC %process 
in

@@

K^{\mathrm{ab}}_{nm}(-\mathbf{k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k})$,     which

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system of
coordinates along  the polar angle, defined as
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and the azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H$_1$ & -0.615 & -1.774 &  0.731 \\
H$_2$ &  0.615 &  0.355 &  0.731 \\
C$_1$ & -0.615 & -1.772 & -0.491 \\
C$_2$ & -0.615 & -0.356 & -0.723 \\
C$_3$ &  0.615 &  0.357 & -0.490 \\
C$_4$ &  0.615 &  1.774 & -0.731 \\
\hline
\end{tabular}
\caption{
Atomic positions in the unit cell of the Up structure shown
in Fig. \ref{fig:up-struc}. 
}
\label{tab:up-unitcell}
\end{table}
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H$_1$ &  -0.615 &  -1.421 & \ 1.472 \\
C$_1$ &  -0.615 &  -1.733 & \ 0.396 \\
C$_2$ & \ 0.615 & \ 1.733 & \ 0.158 \\
C$_3$ & \ 0.615 & \ 0.422 &  -0.158 \\
C$_4$ &  -0.615 &  -0.373 &  -0.396 \\
H$_2$ &  -0.615 &  -0.685 &  -1.472 \\
\hline
\end{tabular}
\caption{
Atomic positions in the unit cell of the Alt structure
 shown in Fig. \ref{fig:alt-struc}. 
}
\label{tab:alt-unitcell}
\end{table}

@@

We calculated the self-consistent ground state and the Kohn-Sham states within
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used Hartwigsen-
Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ from Eq.
\eqref{eq:mu}.
The convergence parameters for the calculations,  corresponding to the Up and
Alt structures are cutoff energies up to 65\,Ha, resulting in LDA
energy band gaps of 0.084\,eV and 0.718\,eV, respectively, 
and 14452 ${\mathbf k}$ points in the IBZ
where the energy
eigenvalues and matrix elements were calculated;
to
integrate $\mu^{\mathrm{abcd}}(\omega)$ and
$\xi^{\mathrm{a}\mathrm{b}}(\omega)$ the linearized analytic tetrahedron method
(LATM) has been used.\cite{nastosPRB07} 
We neglect the anomalous velocity term
$\hbar(\boldsymbol{\sigma}\times\nabla V)/4m^2c^2$, where $V$ is the crystal
potential, in $\hat{\mathbf v}$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore, $[\hat{\mathbf
v},\hat{\mathbf S}]=0$, and Eq.~\eqref{z.1} reduces to $\hat
K^{\mathrm{a}\mathrm{b}}=\hat v^\mathrm{a} \hat S^\mathrm{b}=\hat S^\mathrm{b}
\hat v^\mathrm{a}$. Finally, the prime in the sum of Eq.~\eqref{eq:mu} is
restricted to quasi-degenerated conduction bands $c$ and $c'$ that are closer 
than 30 meV to each other, which  is both typical laser- pulse energy width and the thermal
room-temperature energy level broadening.\cite{nastosPRB07}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption Spin velocity injection {$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  {\it vs.} photon energy $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The
  largest velocity are at the low energy regions of the spectra for the Alt and
  Up structures, becoming different from zero at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ starts rising  
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$, parallel to the surface, with their spin
polarized along $z$, perpendicular to the surface. Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$
from Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$.
For CdSe, we find that for $\hbar\omega=0.844$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and ${\cal
V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at
$\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal
V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with $\mathrm{a}=x,y,z$.
For these bulk semiconductors, the $x$, $y$, and $z$ axis are taken along the
standard cubic unit cell directions, [100], [010], and [001], respectively. In
Table \ref{tab:vab-str-comp}, we compare ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and
bulk crystals. We stress that, as shown in the figure, for the 2D structures
have maxima in ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ are higher than
for the bulk crystals more than order of magnitude. In particular, the Alt structure demonstrates a ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$  about  12 times larger than that
of CdSe and GaAs.
%%%%%%%%%   #### Finished here
\begin{table}%[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
$Up$  & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
$Alt$ & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and
their corresponding polarization angle $\alpha$ and energy $\hbar\omega$ . }
\label{tab:vab-str-comp}
\end{table}
